Binning Strategies and Related Loss for Graphics of Large Data

### 2013-03-06 Paper.Rnw
Fixed std/random "plaid" graph.

LossTable <- NULL
bins <- rbind(c(2,2), c(5,25))
computeLoss <- function(binning, data){
  lRandom <- lossCalc(data, binning=binning, type="random", newData=TRUE)
  lMean <- lossCalc(data, binning=binning, type="standard", newData=TRUE)
  datplot <- rbind(cbind(lRandom$NewData, Algorithm="Random"), cbind(lMean$NewData, Algorithm="Standard"))
  qplot(G, SO, fill = Freq, geom="tile", data = datplot, 
      main=paste("Baseball: Binning by (", binning[1], ", ", binning[2],")", sep="")) + 
    facet_grid(.~Algorithm) +
    scale_fill_gradient(low="#56B1F7", high="#132B43", guide="legend", trans="log", 
                        breaks=c(1, 5, 50, 500, 5000)) + 
    theme_bw() + theme(legend.position="bottom")
  ggsave(paste("./figure/RandomStd", binning[1], binning[2], ".pdf", sep=""), 
         width=8, height=5, units="in", dpi=1200)

@@

At the example of the pitching data, we will %consider the variables Games and Strikeouts, which are strongly correlated. This example 
highlight an important advantage of random binning. The two tile plots at the top of Figure~\ref{fig:binning-comparison} show bins of the same width, but clear differences between the binning results are evident: while the standard binning algorithm has rounding artifacts that are clearly visible for even bin widths, the random algorithm produces a graph that better represents the true shape of the data.
\begin{figure}[hbtp]
\centering
\subfloat[(2, 2) tiling of G, SO data]{
\includegraphics[keepaspectratio=true, width=.9\textwidth]{images/RandomStd22.pdf}
}\\
\subfloat[(5, 25) tiling of G, SO data]{
\includegraphics[keepaspectratio=true, width=.9\textwidth]{images/RandomStd525.pdf}
}
\caption{\label{fig:binning-comparison} Comparison of random and standard binning: standard binning introduces artificial striping when the chose bin width is not a multiple of the data resolution.}
\end{figure}

>>>>

LossTable <- NULL
bins <- rbind(c(2,10), c(5,25))
computeLoss <- function(binning, data){
  lRandom <- lossCalc(data, binning=binning, type="random", newData=TRUE)
  lMean <- lossCalc(data, binning=binning, type="standard", newData=TRUE)
  datplot <- rbind(cbind(lRandom$NewData, Algorithm="Random"), cbind(lMean$NewData, Algorithm="Standard"))
  datplot <- ddply(datplot, .(G, SO, Algorithm), summarise, Freq=sum(Freq))
  qplot(G, SO, fill = Freq, geom="tile", data = datplot, 
      main=paste("Baseball: Binning by (", binning[1], ", ", binning[2],")", sep="")) + 
    facet_grid(.~Algorithm) +
    scale_fill_gradient(low="#56B1F7", high="#132B43", guide="legend", trans="log", 
                        breaks=c(1, 5, 50, 500, 5000)) + 
    theme_bw() + theme(legend.position="bottom")
  ggsave(paste("./figure/RandomStd", binning[1], binning[2], ".pdf", sep=""), 
         width=8, height=5, units="in", dpi=1200)

@@

At the example of the pitching data, we will %consider the variables Games and Strikeouts, which are strongly correlated. This example 
highlight an important advantage of random binning. The two tile plots at the top of Figure~\ref{fig:binning-comparison} show bins of the same width, but clear differences between the binning results are evident: while the standard binning algorithm has rounding artifacts that are clearly visible for even bin widths, the random algorithm produces a graph that better represents the true shape of the data.
\begin{figure}[hbtp]
\centering
\subfloat[(2, 2) tiling of G, SO data]{
\includegraphics[keepaspectratio=true, width=.9\textwidth]{images/RandomStd210.pdf}
}\\
\subfloat[(5, 25) tiling of G, SO data]{
\includegraphics[keepaspectratio=true, width=.9\textwidth]{images/RandomStd525.pdf}
}
\caption{\label{fig:binning-comparison} Comparison of random and standard binning: standard binning introduces artificial striping when the chose bin width is not a multiple of the data resolution.}
\end{figure}
